---
title: "gtsummary Overview"
author: 'Lupe Antonio Lopez'
output: pdf_document
date: "March 27, 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why {gtsummary}

- It provides publication & presentation -ready analytical and summary tables using R.

- It allows you summarize datasets, regression models, customize tables, and more in order to generate reproducible reports.

- The code to create tables is consise and customizable. 

- Can summarize datasets/dataframes using mean(), median(), min(), max(), etc.



## Installation

```{r, eval=FALSE, message = FALSE}
install.packages('gtsummary')
```

Once installed, load package for use in file:

```{r}
library(gtsummary)
```



## Tutorial


### Load dataset

```{r, message = FALSE}
#other libraries used:
library(tidyverse)


#included in the package:
data("trial")
```

```{r}
#quick view
head(trial)
```

The dataset contains information on 200 patients receiving two types of chemotherapy treatments (Drug A or B).





### Creating summary tables

Within the `gtsummary` package we can create and customize summary tables using `tbl_summary()`.

Below is a basic template:

```{r, eval=FALSE}
tbl_summary(by =,
            statistic = list(all_continuous() ~ '{mean} ({sd})',
                             all_categorical() ~ '{n} / {N} ({p}%)'),
            digits = all_continuous() ~ 3)
```

#### The basics:

- `by`: input the groups to split the data into

- `statistic`: indicate the summary statistics to be generated \newline
               differs by data type (i.e., continuous, categorical, etc.)
               
- `digits`: number of digits to appear in table \newline
            can customize by variables



some other modifiable input options include:

- `label`: specify variable names to be presented

- `missing`: whether to display row with number of missing observations

- `missing_text`: text label for the missing number row

- `include`: list of variables to include in summary table \newline
             (can also do this beforehand with `select()`)



**Important**: use `list()` to pass more than 2 sets of instructions to each input



#### Functions to add information and format tables

*added using pipes after tbl_summary()*

- `add_p()`: adds p-values to table comparing values across groups
(detects variable type and uses appropriate statistical tests)

- `add_overall()`: adds a column with overall summary statistics

- `add_n()` adds a column with N for non-missing observations for each variable

- `add_stat_label()`: adds label for summary statistics shown in each row

- `modify_header()`: updates column headers

- `bold/italicize_labels()`: bolds/italicizes variable labels within table

- `bold/italicize_levels()`: bolds/italicizes variable levels

- `modify_caption()`: updates table caption/title

- `modify_spanning_header()`: adds/updates a spanning header (i.e., btwn 2 groups)

- `modify_footnote()`: updates footnote






### Using regression models

Within the `gtsummary` package we can create and customize tables from model outputs using `tbl_regression()`.

Below is a basic template:


```{r, message = FALSE}
#fit model:
mod <- glm(response ~ trt + age + grade + stage, data = trial,
           family = binomial(link = 'logit'))


#table:
 tbl_name <- tbl_regression(mod, exponentiate = TRUE)
```

can add functions: 

- `add_global_p()`: adds global p-values

- `add_glance_table()`: adds various model statistics

... and many more.





### Reporting results in-line:

`gtsummary` allows for results from tables to be reported efficiently using `inline_text()`. The default pattern within this function is: `{estimate} ({conf.level*100}% CI {conf.low}, {conf.high}; {p.value})`.


### Example:

*Code*: The odds ratio for age is `r inline_text(tbl_name, age)`




### Examples


```{r}
trial %>% 
  tbl_summary()
```


```{r}
trial %>%
  tbl_summary(
    include = c(trt, age, grade, marker),
    by = trt,
    missing = 'no',
    statistic = list(age ~ '{mean}',
                     c(marker) ~ '{mean} ({sd})',
                     all_categorical() ~ '{n} / {N} ({p}%)'),
    digits = list(all_continuous() ~ 3,
                  c(grade) ~ 2),
    label = grade ~ 'Tumor Grade')
```



```{r, message=FALSE}
#adding in other functions
trial %>%
  tbl_summary(
    include = c(trt, age, grade, marker),
    by = trt,
    missing = 'no',
    statistic = list(c(age) ~ '{mean}',
                     c(marker) ~ '{mean} ({sd})',
                     all_categorical() ~ '{n} / {N} ({p}%)'),
    digits = list(all_continuous() ~ 3,
                  c(grade) ~ 2),
    label = grade ~ 'Tumor Grade') %>%
  add_p() %>%
  add_n() %>%
  add_overall() %>%
  modify_spanning_header(c('stat_1', 'stat_2') ~ '**Treatment**') %>%
  bold_labels() %>%
  modify_header(label = '**Variable**') %>%
  modify_caption('Table 1')

#change header: label = 'Variable'
#bold spanning header
#change caption/title
```










### Additional Resources:

- [https://www.danieldsjoberg.com/gtsummary](https://www.danieldsjoberg.com/gtsummary/#:~:text=The%20%7Bgtsummary%7D%20package%20provides%20an,defaults%20with%20highly%20customizable%20capabilities.)

- https://www.danieldsjoberg.com/gtsummary/articles/tbl_summary.html

- https://www.danieldsjoberg.com/gtsummary-Columbia-Computing-Club/#12



