---
title: "SF tutorial"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(1)
```

# SF package

```{r}
library(tidyverse)

library(sf)

library(sfheaders)
```

sf - basically a tibble, but one of the columns hold geometric / geographical information (like a shape, or point).

sfheaders - helps with transition from sf to tibble, and vice versa.

# Getting data

There are tons of shapefiles online. This is one such [example](https://www2.census.gov/geo/tiger/TIGER2019/ZCTA5/).

```{r}
nyc = read_sf(dsn = "tl_2019_us_zcta510/", layer = 'tl_2019_us_zcta510') %>%
  mutate(borough = case_when(
    ZCTA5CE10 %in% 10001:10282 ~ "Manhattan",
    
    ZCTA5CE10 %in% 10301:10314 ~ "Staten Island",
    
    ZCTA5CE10 %in% 10451:10475 ~ "Bronx",
    
    ZCTA5CE10 %in% c(11004:11109, 11351:11697) ~ "Queens",
    
    ZCTA5CE10 %in% 11201:11256 ~ "Brooklyn",
    
    T ~ "Not NYC"
  )) %>%
  filter(borough != "Not NYC")

head(nyc, 4)
```

# Quick map!

```{r}
ggplot(data = nyc) +
  geom_sf(aes(geometry = geometry), fill = "white")
```

Can also filter, just like a tibble!

```{r}
nyc %>%
  filter(borough != "Bronx") %>%
  ggplot() +
    geom_sf(aes(geometry = geometry))
```

Can also color them by borough!

```{r}
ggplot(data = nyc) +
  geom_sf(aes(geometry = geometry, fill = borough))
```

If I have some sort of numeric measurement (like, percent of population sick in the zip code), we can plot that as well.

```{r}
nyc %>%
  mutate(sick = runif(n = nrow(nyc))) %>%
  ggplot() +
    geom_sf(aes(geometry = geometry, fill = sick))
```

# Center of each zip code

Ideally, we'd want to show the population for each zip code as number in the center of the polygon.

```{r}
nyc = nyc %>%
 mutate(centers = st_centroid(nyc)$geometry,
        people = round(runif(nrow(nyc), max = 100, min = 20)))

head(nyc, 4)

```

```{r}
nyc %>%
  mutate(sick = runif(n = nrow(nyc))) %>%
  filter(borough == "Brooklyn") %>%
  ggplot() +
    geom_sf(aes(geometry = geometry, fill = sick)) +
    geom_sf_text(aes(geometry = centers, label = people), size = 2, color = "orange")
```

We can try fixing the centers so they are *on* the polygon (useful in the case of islands, for example)

```{r}
nyc = nyc %>%
 mutate(centers = st_point_on_surface(nyc)$geometry,
        people = round(runif(nrow(nyc), max = 100, min = 20)))

nyc %>%
  mutate(sick = runif(n = nrow(nyc))) %>%
  filter(borough == "Brooklyn") %>%
  ggplot() +
    geom_sf(aes(geometry = geometry, fill = sick)) +
    geom_sf_text(aes(geometry = centers, label = people), size = 2, color = "orange")
```

# Finding the zip code of points

Sometimes, we may get the specific coordinates and want to know the zip-code/borough/state where the coordinate is on.

```{r}
N <- 100
points =   data.frame(x = rnorm(N, mean = -73.9, sd = 0.2), 
                 y = rnorm(N, mean = 40.7, sd = 0.1))

#source: https://stackoverflow.com/questions/68247417/r-how-to-create-points-using-st-point-applied-to-columns-of-data-frame
points = points %>%
  as.matrix() %>%
  st_multipoint() %>%
  st_sfc() %>%
  st_cast('POINT') %>%
  st_as_sf(crs = st_crs(nyc)) #in this step, we are specifying the 
                              #projection used
```

Let's plot it:

```{r}
ggplot() +
  geom_sf(data = nyc, aes(geometry = geometry), fill = "white") +
  geom_sf(data = points)
```

Now, we can figure out the zip code and borough for each of the points:

```{r}
points = points %>%
          st_join(nyc, join = st_within) %>%
          select(ZCTA5CE10, borough, geometry)

points
```

# Bonus: sampling from geometries

```{r}
points = st_sample(nyc, size = 1000)

ggplot() +
  geom_sf(data = nyc, aes(geometry = geometry), fill = "white") +
  geom_sf(data = points)
```
